Self-identification of gender and sexual orientation in the US General Social Survey:

### 2022-03-28 scripts/scratch.rmd
Reshuffle folders + README

```{r}
gss_clean <- gss_all |>
  mutate(
    id = paste0(year, '_', id),
    sexnow = coalesce(sexnow, sexnow1),
    sexbirth = coalesce(sexbirth, sexbirth1),
    ) |>
  select(id, year, sex, sexnow, sexbirth, sexornt, sexsex, sexsex5)
```
```{r}

@@

```{r}
gss_clean |>
  filter(!if_any(c(sexsex, sexnow, sexornt), is.na)) |>
  ggplot(aes(x = as_factor(sexsex), fill=as_factor(sexsex))) +
  geom_histogram(stat="count") +
  scale_fill_manual(values=c('blue', 'purple', 'red')) +
  theme(axis.text.x = element_blank()) +
  facet_grid(sexnow ~ sexornt,
             scales = "free_y")
```
```{r}
gss_clean |>
  filter(!if_any(c(sexsex5, sexnow, sexornt), is.na)) |>
  ggplot(aes(x = as_factor(sexsex5), fill=as_factor(sexsex5))) +
  geom_histogram(stat="count") +
  scale_fill_manual(values=c('blue', 'purple', 'red')) +
  #geom_text(stat="count", aes(label=..count..), position = position_stack(vjust=.5), color="black") +
  theme_bw() + 
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.major.x = element_blank()
    ) +
  facet_grid(sexornt ~ sexnow,
             scales = "free_y",
             labeller = label_wrap_gen(width = 16, multi_line = T)
             )
```

@@

>>>>

```{r}
gss_clean <- gss_all |>
  mutate(
    id = paste0(year, '_', id),
    sexnow = coalesce(sexnow, sexnow1),
    sexbirth = coalesce(sexbirth, sexbirth1),
    )# |>
  #select(id, year, sex, sexnow, sexbirth, sexornt, sexsex, sexsex5, ptnrornt)
```
```{r}

@@

```{r}
subsetWith(c('sexsex', 'sexnow', 'sexornt')) |>
  ggplot(aes(x = as_factor(sexsex), fill=as_factor(sexsex))) +
  geom_histogram(stat="count") +
  scale_fill_manual(values=c('blue', 'purple', 'red')) +
  theme(axis.text.x = element_blank()) +
  facet_grid(as_factor(sexnow) ~ as_factor(sexornt),
             scales = "free_y",
             labeller = label_wrap_gen(width = 16, multi_line = T)
             )
```
```{r}
gss_clean |>
  filter(!if_any(c(sexsex5, sexnow, sexornt), is.na)) |>
  ggplot(aes(x = as_factor(sexsex5), fill=as_factor(sexsex5))) +
  geom_histogram(stat="count") +
  scale_fill_manual(values=c('blue', 'purple', 'red')) +
  #geom_text(stat="count", aes(label=..count..), position = position_stack(vjust=.5), color="black") +
  theme_bw() + 
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.major.x = element_blank()
    ) +
  facet_grid(as_factor(sexornt) ~ as_factor(sexnow),
             scales = "free_y",
             labeller = label_wrap_gen(width = 16, multi_line = T)
             )
```

@@

```{r}
sex_codes <- c('M', 'F', 'I')
gender_codes <- c('M', 'F', 'T', 'X')

subsetWith(c('sex', 'sexbirth', 'sexnow')) |>
  mutate(
    code = paste0(sex_codes[sexbirth], gender_codes[sexnow])
    ) |>
  select(code, sex) |>
  group_by(code) |>
  table() |>
  knitr::kable(col.names = c("Coded male", "Coded female")) |>
  kableExtra::kable_classic()
```

```{r}
subsetWith(c('sexornt', 'ptnrornt')) |>
  select('sexornt', 'ptnrornt') |>
  filter(ptnrornt != 4) |> # 'I've never had a sexual partner'
  group_by(sexornt) |>
  ggplot(aes(x = as_factor(sexornt), y = as_factor(ptnrornt))) + 
  geom_bin_2d() +
  scale_fill_viridis_c(option="C")
```

```{r}
subsetWith(c('sexnow', 'sexornt', 'ptnrornt', 'ptnrsxnow')) |>
  select('sexnow', 'sexornt', 'ptnrornt', 'ptnrsxnow') |>
  ggplot(aes(x = as_factor(ptnrornt), fill = as_factor(ptnrsxnow))) +
  geom_histogram(stat='count') +
  theme_bw() +
  theme(axis.text.x = element_blank()) +
  labs(
    title = "Gender identity of subject",
    x = "Sexual orientation of last sex partner (Gay, Bisexual, Straight)",
    y = "Sexual orientation of subject",
    fill = "Gender identity of last sex partner"
  ) +
  facet_grid(as_factor(sexornt) ~ as_factor(sexnow), 
  scale="free_y",          
  labeller = label_wrap_gen(width = 16, multi_line = T))
```

```{r}
# based on `sex`, how many respondents have primarily been having
# sexual relationships with an unexpected cohort? (e.g., straight men primarily having sex with men)
subsetWith('sex', 'sexornt', 'sexsex') |>
  select(sex, sexornt, sexsex) |>
  mutate(
    expected_sexsex = case_when(
      sex == 1 & sexornt == 3 ~ 3,
      sex == 1 & sexornt == 1 ~ 1,
      sex == 2 & sexornt == 3 ~ 1,
      sex == 2 & sexornt == 1 ~ 3,
    )
  ) |>
  filter(
    #!is.na(expected_sexsex) &
    expected_sexsex != sexsex
  )
```
```{r}
# sim. with sexnow and sexsex5
subsetWith('sexnow', 'sexornt', 'sexsex5') |>
  select(year, sexnow, sexornt, sexsex5) |>
  rowwise() |>
  mutate(
    label = paste(
      c("Gay", "Bisexual", "Straight")[sexornt],
      c("men", "women", "transgender people", "people of other genders")[sexnow],
      "primarily having sex with",
      c("men", "women", "both men and women")[sexsex5]
    , collapse = " "),
    expected_sexsex = case_when(
      sexnow == 1 & sexornt == 3 ~ 3,
      sexnow == 1 & sexornt == 1 ~ 1,
      sexnow == 2 & sexornt == 3 ~ 1,
      sexnow == 2 & sexornt == 1 ~ 3,
    )
  ) |>
  count(label)

# |>
#   filter(
#     #!is.na(expected_sexsex) &
#     sexsex != expected_sexsex
#   )
```
